---
title: "Take-home Exercise 4"
description: |
  A new article created using the Distill format.
author:
  - name: LIU Zhenglin
    url: https://example.com/norajones
    affiliation: SMU SCIS
    affiliation_url: https://example.com/spacelysprokets
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE)
```


# Getting started

```{r}
packages = c('scales', 'viridis', 
             'lubridate', 'ggthemes', 
             'gridExtra', 'tidyverse', 
             'readxl', 'knitr',
             'data.table', 'ViSiElse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```
# Importing Data

### Read All Activity Logs Files In A list

```{r, echo=TRUE, eval=FALSE}
files <- list.files('E:/data/Activity Logs/')
```

### Find ID of Participants According to Joviality

```{r, echo=TRUE, eval=FALSE}
part <- read_csv("E:/data/Attributes/Participants.csv")
p1 <- part %>% 
  filter(joviality == max(part$joviality))
p2 <- part %>% 
  filter(joviality == min(part$joviality))
p1
p2
```

### Extract Data From Activity Logs of Participants

```{r, eval=FALSE, echo=FALSE}
parts<- c(p1$participantId, p2$participantId)
df <- data.frame()
for (i in files) {
  psl <- read_csv(paste0('E:/data/Activity Logs/', i)) %>%
    filter(participantId %in% parts)
  df <- rbind(df, psl)
}
write_rds(df, "D:/ZhenglinLIU/ISSS_608/th_ex/data/participant_logs.rds")
```

### Importing Data from RDS file

```{r}
logs <- read_rds("data/participant_logs.rds")
logs_1 <-  logs %>% 
  filter(participantId == 113)
logs_2 <-  logs %>% 
  filter(participantId == 758)
```

The participant whose participantId is 113 have the highest joviality
The participant whose participantId is 758 have the lowest joviality


# Data Preparation


```{r}
log_t <- logs %>% 
  mutate(date = date(timestamp)) %>% 
  mutate(month = month(timestamp)) %>% 
  mutate(week = week(timestamp)) %>% 
  mutate(weekday = weekdays(timestamp)) %>% 
  mutate(hour = hour(timestamp)) %>% 
  mutate(min = minute(timestamp)) %>% 
  mutate(time_line = hour*60+min) %>% 
  mutate(day = as.integer(date-min(date))+1)
```


### Change the dataframe into the structure can be used in Visielse

```{r}
hunger113_w <- log_t %>% 
  filter(participantId == 113) %>% 
  filter(weekday %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) %>% 
  select(participantId, hungerStatus, day, time_line) %>% 
  group_by(hungerStatus,day) %>%
  summarise(value = min(time_line)) %>% 
  pivot_wider(day,names_from = hungerStatus,values_from =value)

hunger113_r <- log_t %>% 
  filter(participantId == 113) %>% 
  filter(weekday %in% c("Saturday", "Sunday")) %>% 
  select(participantId, hungerStatus, day, time_line) %>% 
  group_by(hungerStatus,day) %>%
  summarise(value = min(time_line)) %>% 
  pivot_wider(day,names_from = hungerStatus,values_from =value)

hunger758_w <- log_t %>% 
  filter(participantId == 758) %>% 
  filter(weekday %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) %>% 
  select(participantId, hungerStatus, day, time_line) %>% 
  group_by(hungerStatus,day) %>%
  summarise(value = min(time_line)) %>% 
  pivot_wider(day,names_from = hungerStatus,values_from =value)

hunger758_r <- log_t %>% 
  filter(participantId == 758) %>% 
  filter(weekday %in% c("Saturday", "Sunday")) %>% 
  select(participantId, hungerStatus, day, time_line) %>% 
  group_by(hungerStatus,day) %>%
  summarise(value = min(time_line)) %>% 
  pivot_wider(day,names_from = hungerStatus,values_from =value)
```

```{r}
sleep113_w <- log_t %>% 
  filter(participantId == 113) %>% 
  filter(weekday %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) %>% 
  select(participantId, sleepStatus, day, time_line) %>% 
  group_by(sleepStatus,day) %>%
  summarise(value = min(time_line)) %>% 
  pivot_wider(day,names_from = sleepStatus,values_from =value)

sleep113_r <- log_t %>% 
  filter(participantId == 113) %>% 
  filter(weekday %in% c("Saturday", "Sunday")) %>% 
  select(participantId, sleepStatus, day, time_line) %>% 
  group_by(sleepStatus,day) %>%
  summarise(value = min(time_line)) %>% 
  pivot_wider(day,names_from = sleepStatus,values_from =value)

sleep758_w <- log_t %>% 
  filter(participantId == 758) %>% 
  filter(weekday %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) %>% 
  select(participantId, sleepStatus, day, time_line) %>% 
  group_by(sleepStatus,day) %>%
  summarise(value = min(time_line)) %>% 
  pivot_wider(day,names_from = sleepStatus,values_from =value)

sleep758_r <- log_t %>% 
  filter(participantId == 758) %>% 
  filter(weekday %in% c("Saturday", "Sunday")) %>% 
  select(participantId, sleepStatus, day, time_line) %>% 
  group_by(sleepStatus,day) %>%
  summarise(value = min(time_line)) %>% 
  pivot_wider(day,names_from = sleepStatus,values_from =value)

```


```{r}
mode113_w <- log_t %>% 
  filter(participantId == 113) %>% 
  filter(weekday %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) %>% 
  select(participantId, currentMode, day, time_line) %>% 
  group_by(currentMode,day) %>%
  summarise(value = min(time_line)) %>% 
  pivot_wider(day,names_from = currentMode,values_from =value)

mode113_r <- log_t %>% 
  filter(participantId == 113) %>% 
  filter(weekday %in% c("Saturday", "Sunday")) %>% 
  select(participantId, currentMode, day, time_line) %>% 
  group_by(currentMode,day) %>%
  summarise(value = min(time_line)) %>% 
  pivot_wider(day,names_from = currentMode,values_from =value)

mode758_w <- log_t %>% 
  filter(participantId == 758) %>% 
  filter(weekday %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) %>% 
  select(participantId, currentMode, day, time_line) %>% 
  group_by(currentMode,day) %>%
  summarise(value = min(time_line)) %>% 
  pivot_wider(day,names_from = currentMode,values_from =value)

mode758_r <- log_t %>% 
  filter(participantId == 758) %>% 
  filter(weekday %in% c("Saturday", "Sunday")) %>% 
  select(participantId, currentMode, day, time_line) %>% 
  group_by(currentMode,day) %>%
  summarise(value = min(time_line)) %>% 
  pivot_wider(day,names_from = currentMode,values_from =value)
```


Participant 113 weekdays

```{r}
w_113 <- mode113_w %>% 
  merge(hunger113_w, by = "day") %>% 
  merge(sleep113_w, by = "day")
p113_w <- visielse(w_113,doplot = F)
plot(p113_w, 
     vp0w = 0.7, 
     unit.tps = "min", 
     scal.unit.tps = 30, 
     main = "Participant 113 weekdays")
```

Participant 113 weekend

```{r}
r_113 <- mode113_r %>% 
  merge(hunger113_r, by = "day") %>% 
  merge(sleep113_r, by = "day")
p113_r <- visielse(r_113,doplot = F)
plot(p113_r, 
     vp0w = 0.7, 
     unit.tps = "min", 
     scal.unit.tps = 30, 
     main = "Participant 113 weekend")
```


Participant 758 weekdays

```{r}
w_758 <- mode758_w %>% 
  merge(hunger113_w, by = "day") %>% 
  merge(sleep113_w, by = "day")
p758_w <- visielse(w_758,doplot = F)
plot(p758_w, 
     vp0w = 0.7, 
     unit.tps = "min", 
     scal.unit.tps = 30, 
     main = "Participant 758 weekdays")
```

Participant 758 weekend

```{r}
r_758 <- mode758_r %>% 
  merge(hunger113_r, by = "day") %>% 
  merge(sleep113_r, by = "day")
p758_r <- visielse(r_758,doplot = F)
plot(p758_r, 
     vp0w = 0.7, 
     unit.tps = "min", 
     scal.unit.tps = 30, 
     main = "Participant 758 weekend")
```